---
title: "PROVISIONAL. Spatial spread of a bark beetle outbreak in British Columbia"
subtitle: "PLACEHOLDER TITLE. Not chosen by Seamus. Titles and headings are fixed once he chooses them; nothing here binds."
date: today
author:
  - name: Seamus Murphy
    orcid: 0000-0002-1792-0351
abstract: >
  Pending. The framing gate is open and blocking; no abstract is written until the
  question is approved, the prior-work survey has run, the analysis is designed and the
  pre-registration is frozen. This shell exists so that the manuscript's structure,
  formatting apparatus and pre-registered table shells are in place before any outcome
  is computed, exactly as in Study 1.
keywords:
  - bark beetle outbreaks
  - spatial spread
  - travelling wave
  - spatial synchrony
  - British Columbia
format:
  docx:
    prefer-html: true
    reference-doc: ../04.references/style-formal.docx
    highlight-style: pygments
  html:
    toc: true
    number-sections: false
    embed-resources: true
    code-fold: false
    df-print: kable
execute:
  eval: true
  echo: true
  warning: false
  message: false
  error: true
  comment: NA
engine: knitr
citeproc: true
editor: visual
bibliography: ../04.references/references.bib
csl: ../04.references/apa.csl
---

```{r}
#| label: setup
#| echo: true
#| output: false
library(sf)
library(terra)
library(flextable)

root <- normalizePath("..")
inputs <- file.path(root, "02.inputs")
outputs <- file.path(root, "03.outputs")
derived <- file.path(inputs, "derived")
tables <- file.path(outputs, "tables")

BC_ALBERS <- 3005

# Figures are written to 03.outputs/figures as well as embedded, so the README
# and any submission package reference committed files rather than extracting
# images from a rendered document.
#
# The path must be relative to this document. Pandoc resolves an image src
# against the document's directory, so an absolute path arrives as
# "./Volumes/..." which resolves to nothing, and embed-resources then drops
# every figure without failing the render.
knitr::opts_chunk$set(fig.path = "../03.outputs/figures/",
                      dpi = 300, dev = "png")

# House table format: 9 pt, centred headers, numeric right and text left,
# single line spacing, autofit to full width.
fmt_tbl <- function(df, right = NULL) {
  right_cols <- union(names(df)[vapply(df, is.numeric, TRUE)], right)
  left_cols <- setdiff(names(df), right_cols)
  ft <- flextable::flextable(df)
  ft <- flextable::fontsize(ft, size = 9, part = "all")
  ft <- flextable::align(ft, align = "center", part = "header")
  if (length(right_cols))
    ft <- flextable::align(ft, j = right_cols, align = "right", part = "body")
  if (length(left_cols))
    ft <- flextable::align(ft, j = left_cols, align = "left", part = "body")
  ft <- flextable::line_spacing(ft, space = 1, part = "all")
  ft <- flextable::padding(ft, padding.top = 1, padding.bottom = 1, part = "all")
  flextable::set_table_properties(ft, layout = "autofit", width = 1)
}

# Every number in the prose is read from a saved table at render time. A table
# that does not exist yet reads "pending" rather than failing the render or,
# worse, silently carrying a number from a previous run.
grab <- function(file) {
  p <- file.path(tables, file)
  if (file.exists(p))
    utils::read.csv(p, stringsAsFactors = FALSE, check.names = FALSE)
  else NULL
}
val <- function(df, row_col, row_val, col, digits = 3) {
  if (is.null(df)) return("pending")
  i <- which(df[[row_col]] == row_val)
  if (!length(i)) return("pending")
  v <- df[[col]][i[1]]
  if (is.numeric(v)) formatC(v, format = "f", digits = digits) else as.character(v)
}
num <- function(x, digits = 3) if (is.null(x) || is.na(x)) "pending" else
  formatC(x, format = "f", digits = digits)

# Outcome tables. All NULL until the pre-registration is frozen and the
# pipeline has run; every shell below therefore reads "pending".
front_summ <- grab("front-summary.csv")
epi_tbl    <- grab("epicentres.csv")
geom_tbl   <- grab("patch-geometry.csv")
acorr_tbl  <- grab("spatial-autocorrelation.csv")
hyp_tests  <- grab("hypothesis-tests.csv")
ptest      <- grab("pipeline-test.csv")
```

# Introduction

Pending. Written after the framing gate is approved and the prior-work survey has run.
The introduction positions the study against the travelling-wave literature that
@Aukema_2006 opened for this outbreak, and against the scenario models of
@Perevaryukha_2016 and @Perevaryukha_2019, which do not follow it. Its motivating
observation is Study 1's resolution finding: 65 annual observations resolve only two cycle
lengths at the outbreak scale, an independent estimator resolves none, and the archive is
therefore better suited to asking where an outbreak went than how often it returns. The
spatial questions here require no cycle to be resolved.

## Hypotheses

Pending, from the approved framing document. Stated here in the manuscript before any
outcome is computed, as in Study 1.

# Materials and methods

A note on procedure. Pending: this paragraph records which gates were run and which, if
any, were waived, with the compensating discipline stated. Study 1 ran under a waiver of
the survey and design gates; this study does not inherit that waiver.

## Study system

Pending.

## Study area

Pending.

## Data

Pending. Landsat Collection 2 Level-2, tier and footprint named in the framing document,
with the survey polygon archive as the reference tier.

## Analysis window

Pending.

## Front estimation

Pending.

## Geometry statistics

Pending.

## Hypothesis tests

Pending.

## Pipeline test

Pending. Mandatory, and for a spread-front estimator it must include both a synthetic
travelling wave of known speed and direction, which the estimator must recover, and a
synthetic simultaneous eruption, which the estimator must not report as a wave.

# Results

## Pipeline test

```{r}
#| label: tbl-ptest
#| echo: true
#| tbl-cap: "Synthetic pipeline test. Pre-registered shell; reads pending until the test has run."
if (is.null(ptest)) {
  ptest <- data.frame(Scenario = c("planted travelling wave",
                                   "simultaneous eruption",
                                   "spatial noise"),
                      Truth = "pending", Recovered = "pending",
                      Verdict = "pending", check.names = FALSE)
}
fmt_tbl(ptest)
```

## Epicentres

```{r}
#| label: tbl-epicentres
#| echo: true
#| tbl-cap: "Outbreak epicentres: location, first year of detected damage, and the rule under which each was identified. Pre-registered shell; reads pending until outcomes are computed."
if (is.null(epi_tbl)) {
  epi_tbl <- data.frame(Epicentre = "pending", Easting = "pending",
                        Northing = "pending", `First year` = "pending",
                        Rule = "pending", check.names = FALSE)
}
fmt_tbl(epi_tbl)
```

## Spread fronts

```{r}
#| label: tbl-fronts
#| echo: true
#| tbl-cap: "Spread-front speed and direction per epicentre, with the interval method and the number of resampling draws fixed at pre-registration. Pre-registered shell; reads pending until outcomes are computed."
if (is.null(front_summ)) {
  front_summ <- data.frame(Epicentre = "pending",
                           `Speed (km/yr)` = "pending",
                           `Interval` = "pending",
                           `Direction (deg)` = "pending",
                           check.names = FALSE)
}
fmt_tbl(front_summ)
```

## Patch geometry

```{r}
#| label: tbl-geometry
#| echo: true
#| tbl-cap: "Geometry of the annual affected area: area, edge length, shape index and orientation. Pre-registered shell; reads pending until outcomes are computed."
if (is.null(geom_tbl)) {
  geom_tbl <- data.frame(Year = "pending", `Area (km2)` = "pending",
                         `Edge (km)` = "pending", `Shape index` = "pending",
                         check.names = FALSE)
}
fmt_tbl(geom_tbl)
```

## Spatial autocorrelation

```{r}
#| label: tbl-autocorrelation
#| echo: true
#| tbl-cap: "Spatial autocorrelation of annual damage over distance, with the lag structure and the significance procedure fixed at pre-registration. Pre-registered shell; reads pending until outcomes are computed."
if (is.null(acorr_tbl)) {
  acorr_tbl <- data.frame(`Lag (km)` = "pending", Statistic = "pending",
                          p = "pending", check.names = FALSE)
}
fmt_tbl(acorr_tbl)
```

## Hypothesis tests

```{r}
#| label: tbl-tests
#| echo: true
#| tbl-cap: "Hypothesis tests under the pre-stated rules. Pre-registered shell; reads pending until outcomes are computed. The confirmatory family and its multiplicity correction are fixed at pre-registration."
if (is.null(hyp_tests)) {
  hyp_tests <- data.frame(Hypothesis = "pending", Estimate = "pending",
                          p = "pending", `p (Holm)` = "pending",
                          Verdict = "pending", check.names = FALSE)
}
fmt_tbl(hyp_tests)
```

# Discussion and conclusions

Pending. Written as branches before the numbers exist, one branch per possible verdict, so
that the reading of each outcome is fixed in advance, exactly as in Study 1.

## Limitations

Pending.

## Conclusion

Pending.

# Data availability

Pending. All data are expected to be public: Landsat from USGS, the survey polygon archive
and the region boundaries under the Open Government Licence - British Columbia. Every
licence is read from its source on the retrieval date and recorded in the input manifest
before it is stated here.

# Analysis pipeline

Empty. Every chunk in this section is guarded on its own output, so a render executes none
of them, and each is wrapped in `local()` so that a pipeline object cannot replace one the
document is using. Chunks are labelled `pipe-*` and written out to `05.scripts/` by the
`pipe-purl` chunk below. No pipeline chunk is written until the pre-registration is frozen.

```{r}
#| label: pipe-purl
#| echo: true
#| output: false
# Writes the Analysis pipeline chunks out to 05.scripts/ so that long jobs can
# be run from a terminal without rendering the whole document. Files carry a
# do-not-edit header; edits made there are lost on the next regeneration.
local({
  src <- knitr::current_input()
  if (is.null(src) || !file.exists(src)) return(invisible(NULL))
  txt <- readLines(src)
  starts <- grep("^```\\{r\\}", txt)
  labs <- vapply(starts, function(i) {
    j <- i + which(grepl("^#\\| label:", txt[(i + 1):min(i + 6, length(txt))]))[1]
    if (is.na(j)) "" else sub(".*label:\\s*", "", txt[j])
  }, "")
  keep <- grepl("^pipe-", labs) & labs != "pipe-purl"
  for (k in which(keep)) {
    i <- starts[k]
    e <- i + which(txt[(i + 1):length(txt)] == "```")[1]
    body <- txt[(i + 1):(e - 1)]
    body <- body[!grepl("^#\\|", body)]
    f <- file.path(root, "05.scripts", paste0(labs[k], ".R"))
    writeLines(c("# GENERATED FILE. Do not edit.",
                 "# Written from 01.manuscript/manuscript.qmd by the pipe-purl chunk.",
                 "# Edit the chunk in the manuscript and regenerate; edits here are lost.",
                 "# Run from the repository root.", "", body), f)
  }
})
```

# References

::: {#refs}
:::
